---
title: "Zero Variance Plankton Class Differences"
author: "Earl F Glynn"
output: html_document
---

*****

## Invariant features by plankton class?

Nearly 3000 wnd-chrm features were computed for all plankton images.

In the training set, can we find a set of the wnd-chrm features by plankton class that have very low variance?  That is, can we find a set of signature features for each class that are invariant, or nearly invariant?  

Is the set of features common to a class unique among other plankton classes?


```{r}
#setwd("C:/Kaggle/2015/Plankton/Wnd-Charm/")
time.1 <- Sys.time()
format(time.1, "%Y-%m-%d-%H%M%S")
```                                       

### Load scaled TRAINING file 

Also, read *plankton* classifications

```{r, comment=NA}
load("plankton-train-wndchrm-scaled.Rdata", verbose=TRUE) 
```

```{r}
zeroList <- list()
```

### Form list of low (zero) variance features by class

In preliminary experiments, a number of features had zero variance within a class.  Can we use class zero variance features as a signature?

```{r}
for (i in 1:length(unique(plankton)))
{
  feature <- scaledTrain[plankton == i,]
  sigma   <- apply(feature, 2, sd)
 
  zeroList[[i]] <- list()
  zeroList[[i]]$feature   <- feature
  zeroList[[i]]$zeroCount <- sum(sigma == 0)
  zeroList[[i]]$zeroIndex <- which(sigma == 0)
  zeroList[[i]]$values    <- feature[1,which(sigma == 0)] 
}
```

### Explore list of lists

Let's look at the first three classes, which are at least somewhat similar:

#### 1. acantharia_protist (889 images)

```{r, comment=NA}
dim(zeroList[[1]]$feature)
zeroList[[1]]$zeroCount

options(width=140)
data.frame(index=zeroList[[1]]$zeroIndex, value=zeroList[[1]]$values)
```

#### 2. acantharia_protist_big_center (13 images)

```{r, comment=NA}
dim(zeroList[[2]]$feature)
zeroList[[2]]$zeroCount
data.frame(index=zeroList[[2]]$zeroIndex, value=zeroList[[2]]$values)
```

#### 3. acantharia_protist_halo (71 images)

```{r, comment=NA}
dim(zeroList[[3]]$feature)
zeroList[[3]]$zeroCount
data.frame(index=zeroList[[3]]$zeroIndex, value=zeroList[[3]]$values)
```

#### Venn diagram of features in Plankton classes 1, 2, 3 (protists)

```{r, message=NA}

acantharia <- as.integer(zeroList[[1]]$zeroIndex)
bigcenter  <- as.integer(zeroList[[2]]$zeroIndex)
halo       <- as.integer(zeroList[[3]]$zeroIndex)

library(VennDiagram)
venn.plot <- venn.diagram(list(A=acantharia, B=bigcenter, C=halo), 
                          filename="Protists123.tif", width=600, height=600, resolution=100,
                          main="Protists Plankton Common Features",
                          category.names=c("acantharia protist", 
                                           "acantharia protist big center", 
                                           "acantharia protist halo"))
```

![Protists123](Protists123.tif)

#### Set differences

```{r, message=NA}
setdiff(acantharia, bigcenter)
setdiff(acantharia, halo)
setdiff(bigcenter, halo)
```


#### Venn diagram of features in Plankton classes 83 - 87 (protists)

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[83]]$zeroIndex), 
                               B=as.integer(zeroList[[84]]$zeroIndex),
                               C=as.integer(zeroList[[85]]$zeroIndex),
                               D=as.integer(zeroList[[86]]$zeroIndex),
                               E=as.integer(zeroList[[87]]$zeroIndex)),
                          filename="Protists83-87.tif", width=600, height=600, resolution=100,
                          main="Protists Plankton Common Features",
                          category.names=c("dark center", 
                                           "fuzzy olive", 
                                           "noctiluca", 
                                           "other", 
                                           "star"))
```

![Protists8387](Protists83-87.tif)


#### Venn diagram of features in Plankton classes 5 - 8 (appendicularian)

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[5]]$zeroIndex), 
                               B=as.integer(zeroList[[6]]$zeroIndex),
                               C=as.integer(zeroList[[7]]$zeroIndex),
                               D=as.integer(zeroList[[8]]$zeroIndex)),
                          filename="Appendicularian5-8.tif", width=600, height=600, resolution=100,
                          main="Appendicularian Plankton Common Features",
                          category.names=c("fritillaridae", 
                                           "s shape",                         
                                           "slight curve", 
                                           "straight"))
```

![Appendicularian58](Appendicularian5-8.tif)

#### Venn diagram of features in Plankton classes:  Detritus

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[ 50]]$zeroIndex), 
                               B=as.integer(zeroList[[ 34]]$zeroIndex),
                               C=as.integer(zeroList[[ 35]]$zeroIndex),
                               D=as.integer(zeroList[[ 36]]$zeroIndex)),
                          filename="Detrius.tif", width=600, height=600, resolution=100,
                          main="Detritus Common Features",
                          category.names=c("fecal pellet", 
                                           "detritus blob",                         
                                           "detritus filamentous", 
                                           "detritus other"))
```

![Detrius](Detrius.tif)


#### Venn diagram of features in Plankton classes:  Unknown

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[119]]$zeroIndex),
                               B=as.integer(zeroList[[120]]$zeroIndex),                      
                               C=as.integer(zeroList[[121]]$zeroIndex)),
                          filename="Unknown.tif", width=600, height=600, resolution=100,
                          main="Unknown Plankton Common Features",
                          category.names=c("unknown blobs",                                                                  
                                           "unknown sticks", 
                                           "unknown unclassified"))
```

![Unknown](Unknown.tif)

#### Venn diagram of features in Plankton classes:  Fish Larvae

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[51]]$zeroIndex), 
                               B=as.integer(zeroList[[52]]$zeroIndex),
                               C=as.integer(zeroList[[53]]$zeroIndex),  # Leave out 54 for now
                               D=as.integer(zeroList[[55]]$zeroIndex),
                               E=as.integer(zeroList[[56]]$zeroIndex)),
                          filename="FishLarvae51-56.tif", width=600, height=600, resolution=100,
                          main="Fish Larvae Plankton Common Features",
                          category.names=c("deep body", 
                                           "leptocephali",                         
                                           "medium body",                                                                   
                                           "thin body", 
                                           "very thin body"))
```

![FishLarvae5156](FishLarvae51-56.tif)

#### Venn diagram of features in Plankton classes:  Ctenophore

```{r, message=NA}
venn.plot <- venn.diagram(list(A=as.integer(zeroList[[29]]$zeroIndex), 
                               B=as.integer(zeroList[[30]]$zeroIndex),
                               C=as.integer(zeroList[[31]]$zeroIndex),
                               D=as.integer(zeroList[[32]]$zeroIndex)),
                          filename="Ctenophore.tif", width=600, height=600, resolution=100,
                          main="Ctenophore Common Features",
                          category.names=c("cestid", 
                                           "cydippid no tentacles",                         
                                           "cydippid tentacles", 
                                           "lobate"))
```

![Ctenophore](Ctenophore.tif)

*****

```{r}
time.2 <- Sys.time()
cat(sprintf("%.1f", as.numeric(difftime(time.2, time.1, units="secs"))), " secs\n")
```

*****

*efg* @EarlGlynn

`r format(Sys.time(), "%Y-%m-%d  %H%M")`
                                                                          