---
title: "Plankton TRAINING file with wndchrm features"
author: "Earl F Glynn"
output: html_document
---

*****

Load wndchrm feature files for TRAINING cases and save to .Rdata file.

efg, 2015-02-09

```{r}
time.1 <- Sys.time()
format(time.1, "%Y-%m-%d-%H%M%S")

TRAIN.BASE   <- "train"
```

## Parse wndchrm feature file

```{r}
parseWndchrm <- function(wndchrm)
{
  # For now, be very strict with header
  header   <- wndchrm[1]

  filename <- wndchrm[2]

  wndchrm <- wndchrm[-1:-2]
  splits <- strsplit(wndchrm, "\t")

  values   <- as.numeric(unlist(lapply(splits, "[", 1)))
  features <- unlist(lapply(splits, "[", 2))

  invisible( list(header=header, filename=filename, values=values, features=features) )
}
```

## Count files in all TRAIN.BASE directories

```{r}
dirs <- list.dirs(path=TRAIN.BASE, full.names=FALSE, recursive=FALSE)

file_counts <- integer(length(dirs))
for (k in 1:length(dirs))
{
  directory <- dirs[k]
  files <- list.files(path=paste0(TRAIN.BASE, "/", directory),
                      pattern="^.*\\.sig$")
  file_counts[k] <- length(files)
}

Nfiles <- sum(file_counts)
stopifnot(Nfiles == 30336)  # Make sure all files counted
```

## Create empty feature.matrix

Setup feature matrix based on first .sig file in first directory

```{r}
directory <- dirs[1]
files <- list.files(path=paste0(TRAIN.BASE, "/", directory),
                    pattern="^.*\\.sig$")
```

Read wndchrm file but ignore first two rows for now.

```{r}
wndchrm <- readLines(paste0(TRAIN.BASE, "/", directory, "/", files[1]))
parsed <- parseWndchrm(wndchrm)
```

Allocate feature matrix:

* One row for each file.  Once column for each feature, plus
* a column for the plankton class, which will be column 1.

```{r}
feature.matrix <- matrix(0, nrow=Nfiles, ncol=1+length(parsed$features))
dim(feature.matrix)
object.size(feature.matrix)

rowNames    <- rep("", Nfiles)
columnNames <- c("class", parsed$features) # R accepts almost any string as column name
# set dimnames below after read loop when filenames are known
```

## Loop through directories and files to fill feature matrix

```{r}
rowIndex <- 1
for (k in 1:length(dirs))
{
  directory <- dirs[k]
  files <- list.files(path=paste0(TRAIN.BASE, "/", directory),
                      pattern="^.*\\.sig$")

  cat(k, directory, length(files), "\n")
  flush.console()

  rowNames[rowIndex:(rowIndex+length(files)-1)] <- files

  for (i in 1:length(files))
  {
    wndchrm <- readLines(paste0(TRAIN.BASE, "/", directory, "/", files[i]))
    parsed <- parseWndchrm(wndchrm)
    stopifnot(all(columnNames[-1] == parsed$features)) # check column names
    feature.matrix[rowIndex,] <- c(k, parsed$values)
    rowIndex <- rowIndex + 1
  }
}

dimnames(feature.matrix) = list(rowNames, columnNames)
```

Compute table of class IDs to check if files assigned correctly to rows in feature matrix.

```{r}
table(feature.matrix[,1])
```

## Remove constant columns.

Column is marked as constant if column *max* equals *min*.

How is it possible to have so many constant columns?

```{r}
min.col <- apply(feature.matrix, 2, min)
max.col <- apply(feature.matrix, 2, max)
N.removed <- sum(min.col == max.col)

minmax <- which(min.col == max.col)
minmax <- data.frame(ColName=names(minmax), ColNumber=as.integer(minmax))
minmax
write.csv(minmax, "plankton-train-constant-columns.csv")

constant.columns <- feature.matrix[,min.col == max.col]
head(constant.columns, 1)

dim(feature.matrix)
N.removed
feature.matrix <- feature.matrix[,min.col != max.col]
dim(feature.matrix)
```

## Save TRAINING feature.matrix file

```{r}
format(Sys.time(), "%Y-%m-%d-%H%M%S")
save(feature.matrix, file="plankton-train-wndchrm-features.Rdata")
format(Sys.time(), "%Y-%m-%d-%H%M%S")
```

*****

```{r}
time.2 <- Sys.time()
cat(sprintf("%.1f", as.numeric(difftime(time.2, time.1, units="secs"))), " secs\n")
```

*****

*efg* @EarlGlynn

`r format(Sys.time(), "%Y-%m-%d  %H%M")`
