---
title: "Plankton testing file with wndchrm features"
author: "Earl F Glynn"
output: html_document
---

*****

Load wndchrm feature files for test cases and save to .Rdata file.

efg, 2015-02-14

```{r}
time.1 <- Sys.time()
format(time.1, "%Y-%m-%d-%H%M%S")

TEST.BASE   <- "test"
```

## Parse wndchrm feature file

```{r}
parseWndchrm <- function(wndchrm)
{
  # For now, be very strict with header
  header   <- wndchrm[1]

  filename <- wndchrm[2]

  wndchrm <- wndchrm[-1:-2]
  splits <- strsplit(wndchrm, "\t")

  values   <- as.numeric(unlist(lapply(splits, "[", 1)))
  features <- unlist(lapply(splits, "[", 2))

  invisible( list(header=header, filename=filename, values=values, features=features) )
}
```

## Create empty feature.matrix

Setup feature matrix based on first .sig file in directory

```{r}
files <- list.files(path=TEST.BASE, pattern="^.*\\.sig$")
Nfiles <- length(files)
```

Read wndchrm file but ignore first two rows for now.

```{r}
wndchrm <- readLines(paste0(TEST.BASE, "/", files[1]))
parsed <- parseWndchrm(wndchrm)
```

Allocate feature matrix:

* One row for each file.  Once column for each feature.

```{r}
feature.matrix <- matrix(0, nrow=Nfiles, ncol=length(parsed$features))
dim(feature.matrix)
object.size(feature.matrix)

rowNames <- files
columnNames <- parsed$features # R accepts almost any string as column name
```

## Loop through files and fill feature matrix

```{r}
for (i in 1:Nfiles)
{
  if (i %% 5000 == 1)
  {
    cat(i, files[i], "\n")
    flush.console()
  }
  wndchrm <- readLines(paste0(TEST.BASE, "/", files[i]))
  parsed <- parseWndchrm(wndchrm)
  stopifnot(all(columnNames == parsed$features)) # check column names
  feature.matrix[i,] <- parsed$values
}
```

## Remove constant columns.

How is it possible to have so many constant columns?

```{r}
min.col <- apply(feature.matrix, 2, min)
max.col <- apply(feature.matrix, 2, max)
N.removed <- sum(min.col == max.col)

which(min.col == max.col)
write.csv(which(min.col == max.col), "plankton-test-constant-columns.csv")

constant.columns <- feature.matrix[,min.col == max.col]
head(constant.columns, 1)

dim(feature.matrix)
N.removed
feature.matrix <- feature.matrix[,min.col != max.col]
dim(feature.matrix)
```

## Save feature.matrix file

```{r}
format(Sys.time(), "%Y-%m-%d-%H%M%S")
save(feature.matrix, file="plankton-test-wndchrm-features.Rdata")
format(Sys.time(), "%Y-%m-%d-%H%M%S")
```

*****

```{r}
time.2 <- Sys.time()
cat(sprintf("%.1f", as.numeric(difftime(time.2, time.1, units="secs"))), " secs\n")
```

*****

*efg* @EarlGlynn

`r format(Sys.time(), "%Y-%m-%d  %H%M")`

