---
title: "Combine wndchrm and skimage image features"
author: "Earl F Glynn"
---

*****

```{r}
time.1 <- Sys.time()
format(time.1, "%Y-%m-%d-%H%M%S")
```      
# Training set

## Load Training Set files

### Load wndchrm features 

feature.matrix is a matrix

```{r, comment=NA}
load(file="../Wnd-Charm/plankton-train-wndchrm-features.Rdata", verbose=TRUE)   
str(feature.matrix)
trainFeatures <- colnames(feature.matrix)
```

Column 1 of the feature.matrix defines the plankton **class**.

```{r, comment=NA}
counts1 <- table(feature.matrix[,1])
counts1
```

The feature.matrix row names define the image filename.  For example, the original image 100224.jpg was converted to 100224.tif.  The file of wndchrm properties was named 100224-l.sig, which is the rowname of feature.matrix.

### Load skimage features

skimage is a data.frame

```{r, comment=NA}
skimage <- read.csv("../skimage/train-properties-MASTER.csv", as.is=TRUE)
str(skimage)
head(skimage,2)
```

**skimage$train** defines the plankton class.  

**skimage$filename** defines the image filename (e.g., 10024.jpg).

Filename comparisons betweem skimage and the feature.matrix must drop the extension (e.g., 10024).

### Load index conversion chart

The cause was not investigated, but for some reason traversing the directory tree of plankton .tif files using CentOS/IPython resulted in a slightly different directory order than traversing the corresponding directory tree of the original .jpg files.  

Instead of resolving the original problem, a conversion table was created the tells the order for the two sources.

Here skimage rows are ordered to match the feature.matrix rows using filenames.

```{r}
xlate <- read.csv("wndchrm-skimage-index-translate.csv", as.is=TRUE)

skiorder <- skimage$train
```

## Convert skimage class to wndchrm class

Use brute force here

```{r}
for (i in 1:121)
{
  skimage$train[skiorder == xlate$skimage[i]] <- xlate$wndchrm[i]
}
```

### standardize row order sort:  plankton class, filename

```{r}
wndchrm <- feature.matrix[order(feature.matrix[,1], row.names(feature.matrix) ),]
skimage <- skimage[order(skimage$train, skimage$filename), ]
```

## Do classes and files line up?

### Verify match by plankton class

```{r}
all((skimage$train) == feature.matrix[,1])
sum((skimage$train) == feature.matrix[,1])
```

### Verify match by filename

wndchrm
```{r}
splits <- strsplit(row.names(feature.matrix), "-")
wndchrmFileIndex <- unlist(lapply(splits, "[", 1))
head(wndchrmFileIndex)
```

skimage
```{r}
splits <- strsplit(skimage$filename, "\\.")
skimageFileIndex <- unlist(lapply(splits, "[", 1))

all(wndchrmFileIndex == skimageFileIndex)
sum(wndchrmFileIndex == skimageFileIndex)
```

## With files aligned now, combine feature sources

Drop columns skimage$train and skimage$filename since they align with corresponding info in the feature.matrix.

```{r, comment=NA}
str(skimage)

skimage$filename <- NULL
skimage$train    <- NULL

str(skimage)

```

Convert to matrix and cbind to feature.matrix

```{r, comment=NA}
skimage <- as.matrix(skimage)

dim(feature.matrix)
dim(skimage)

feature.matrix <- cbind(feature.matrix, skimage)
dim(feature.matrix)
```

Let's treat the plankton class as a separate vector and remove class from the matrix

```{r}
train.class    <- feature.matrix[, 1]
train.features <- feature.matrix[,-1]
```

Final look at number of images by feature

```{r, comment=NA}
table(train.class)
```

## Save train data to .Rdata file for fast loads
```{r, comment=NA}
dim(train.features)
object.size(train.features)
object.size(train.class)
save(train.features, train.class, file="plankton-train-wndchrm-skimage-features.Rdata")
```

Save training and testing data into separate .Rdata files so testing data is only brought into memory after model has been built.

Save some memory to process testing set now
```{r}
rm(train.features, train.class)
```

# Testing set

## Load Testing Set files

### Load wndchrm features 

```{r, comment=NA}
load(file="../Wnd-Charm/plankton-test-wndchrm-features.Rdata", verbose=TRUE)   
str(feature.matrix)
testFeatures <- colnames(feature.matrix)
```

### Load skimage features

```{r, comment=NA}
skimage <- read.csv("../skimage/test-properties-MASTER.csv", as.is=TRUE)
str(skimage)
head(skimage,2)
```

## Make sure train and test features match.
```{r, comment=NA}
trainFeatures <- trainFeatures[-1]  # drop class from train; not in test

length(trainFeatures)
length(testFeatures)

length(intersect(testFeatures, trainFeatures))  # count of features in common
setdiff(testFeatures, trainFeatures)
setdiff(trainFeatures, testFeatures)
```

### Remove features in test set not in train set

Four features must be removed from the test set.  These features were constant in the training set and were removed, but they were not constant in the test set.

```{r, comment=NA}
featuresToRemove <- setdiff(testFeatures, trainFeatures)
featuresToRemoveIndices <- which(colnames(feature.matrix) %in% featuresToRemove)

colnames(feature.matrix)[featuresToRemoveIndices]

dim(feature.matrix)
feature.matrix <- feature.matrix[,-featuresToRemoveIndices]
dim(feature.matrix)
```

### Final check that wndchrm features in train and test match exactly
```{r, comment=NA}
testFeatures <- colnames(feature.matrix)
length(trainFeatures)
length(testFeatures)
all(testFeatures == trainFeatures)
```

## Make sure wndchrm and skimage files processed in the same order.

### Verify match by filename

wndchrm
```{r}
splits <- strsplit(row.names(feature.matrix), "-")
wndchrmFileIndex <- unlist(lapply(splits, "[", 1))
head(wndchrmFileIndex)
```

skimage
```{r}
splits <- strsplit(skimage$filename, "\\.")
skimageFileIndex <- unlist(lapply(splits, "[", 1))

all(wndchrmFileIndex == skimageFileIndex)
sum(wndchrmFileIndex == skimageFileIndex)
```

## Cleanup on skimage before combining with windchrm features

Drop columns skimage$train and skimage$filename since they align with corresponding info in the feature.matrix.

```{r, comment=NA}
str(skimage)
skimage$filename <- NULL  # no longer needed since aligned
str(skimage)

skimage <- as.matrix(skimage)
```

## cbind feature.matrix (wndchrm) and skimage features

```{r, comment=NA}
dim(feature.matrix)
dim(skimage)

test.features <- cbind(feature.matrix, skimage)
rm(feature.matrix)
```

## Save test data to .Rdata file for fast loads
```{r, comment=NA}
dim(test.features)
object.size(test.features)
save(test.features, file="plankton-test-wndchrm-skimage-features.Rdata")
```

*****

```{r}
time.2 <- Sys.time()
cat(sprintf("%.1f", as.numeric(difftime(time.2, time.1, units="secs"))), " secs\n")
```

*****

*efg* @EarlGlynn

`r format(Sys.time(), "%Y-%m-%d  %H%M")`               